Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-04 methods_paper/fabl-paper.tex
update.

\paragraph{Violations of One-to-One Matching} Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
duplicates found across file for both \texttt{fabl} and \texttt{BRL}. We
see that \texttt{fabl} consistently overmatches within each Gibbs
iteration when compared to \texttt{BRL}, which is to be expected because
\texttt{BRL} explicitly prevents matches that violate one-to-one
matching throughout the entire sampler. Most of such matchings are due
to the randomness in the sampling procedure, and they occur sporadically
throughout the sampler in such a way that does not measurably influence
the eventual Bayes estimate.

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes}

	\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
\end{figure}

Overall, both methods presented similar results. \texttt{fabl} yielded
an initial Bayes estimate of 195 matches found across files, and after
resolving matches that violated one-to-one requirements, yielded 180
matches across files. This is acceptably close to the 181 matches found
by \texttt{BRL}. The main reason for this discrepancy is the difference
in how each model handles situations in which one file in \(B\) has
multiple plausible matches in \(A\).

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records}

	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
\end{figure}

	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing}

	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
\end{figure}

The records shown in Figure~\ref{fig:rosa-maria} provide one such instance. Note these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. Additionally, using the modified Levenstein distance
from Sadinle 2017, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

Figure~\ref{fig:mixing-plot} shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(B\) match to the
same record in \(A\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. Additionally,
\texttt{fabl} allows the modeler to inspect records with multiple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate.

>>>>

\paragraph{Violations of One-to-One Matching}

In this section, we investigate violations of one-to-one matching of our method as compared to \texttt{BRL}.

Figure~\ref{fig:overlap-plot} shows the posterior distribution for \(D\), the number of
duplicates found in both databases for both \texttt{fabl} and \texttt{BRL}. We
see that \texttt{fabl} consistently overmatches within each Gibbs
iteration when compared to \texttt{BRL}, which is to be expected because
\texttt{BRL} explicitly prevents matches that violate one-to-one
matching throughout the entire sampler. Most of such matchings are due
to the randomness in the sampling procedure, and they occur sporadically
throughout the sampler in such a way that does not measurably influence
the eventual Bayes estimate.

\begin{figure}[htbp]
\begin{center}
 \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}
\label{fig:overlap-plot}
\end{center}
\end{figure}

%\begin{figure}[t]
%	
%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/overlap_distribution_smallP_bayes} 
%		
%	}
%	
%	\caption{Posterior distribution of overlap across the two files. The solid lines show the Bayes estimate for the amount of overlap, and the dashed line is the Bayes estimate under fabl before resolving violations of one-to-one matching}\label{fig:overlap-plot}
%\end{figure}

Overall, both methods present similar results. \texttt{fabl} yields
an initial Bayes estimate of total 195 DID. After
resolving matches that violated one-to-one requirements,  we estimate 180 DID. 
\textcolor{red}{Can you report the posterior standard error for each one as we want uncertainty quantification.}

This is very close to the 181 DID found
by \texttt{BRL}. The main reason for this discrepancy is the difference
in how each model handles situations in which one file in \(\bm{X}_1\) has
multiple plausible matches in \(\bm{X}_2\).

%
%\begin{figure}[t]
%	
%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
%		
%	}
%	
%	\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
%\end{figure}

Figure~\ref{fig:rosa-maria} illustrates an example. Note these
records present a near violation of our assumption that there are no
duplications within files; we continue to assume that records 825 and
826 in ERTL correspond to different individuals, but their records are
nearly identical. In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and
\(\gamma_{2776, 826}\) are exactly identical.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/rosa_records} 
\caption{Example of linkage situation with multiple plausible matches}\label{fig:rosa-maria}
\end{center}
\end{figure}

Figure~\ref{fig:mixing-plot} shows the values that \(Z_{825}\) and \(Z_{826}\) take on
throughout the Gibbs sampler, and demonstrates how each method handles
this situation. Under \texttt{fabl}, both records in \(\bm{X}_2\) match to the
same record in \(\bm{X}_1\) throughout the Gibbs process, creating consistent
violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process
creates one matching configuration and stays there for a while. However, if
one pair ``unmatches,'' then the other record has a chance to latch on.
Then, the Gibbs process is stuck with that matching status for a while,
resulting in a Gibbs process with poor mixing. In addition,
\texttt{fabl} allows the modeler to inspect records with multiple
plausible matches, and if they desire, to then choose the record pairing
with the highest posterior probability. \texttt{BRL} in contrast, in
strictly enforcing one-to-one matching throughout the sampler, can lead
to situations where none of the plausible matches reach the threshold to
be identified through the Bayes estimate. In short, we have pointed out why one may not wish to imbed this constraint in a model, and proposed a simple alternative that performs about the same as \texttt{BRL}, while scaling to much larger data sets.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
\end{center}
\end{figure}

%\begin{figure}[t]
%	
%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
%		
%	}
%	
%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
%\end{figure}
